---
title: "Readme"
author: "Cynthia SC"
date: "2025-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

# Habenula multiome

<br>

### Notes:

**R environment**: Make sure to run your script with *R v4.3.x* (not a newer version), as upgrading has been observed to corrupt the chromatin assay in multiome Seurat objects. To avoid unexpected results, ensure you load the appropriate module in your bash scripts:

```bash
# Load the specific R module to safely save/load Seurat v5 chromatin integrity
module load conda_R/4.3.x
```
<br><br>

**clustering directory**: most notable clustering results
```{r}
# Load knitr for kable
library(knitr)
library(here)

# Define the directory path
cluster_dir <- here("processed-data/05_Clustering_ARCr/01_clustering_std_method")
print(cluster_dir)
# List only .rds files
clustering_files <- list.files(path = cluster_dir, pattern = "\\.rds$", full.names = TRUE)

# Create a data frame for display
clustering_df <- data.frame(`RDS File` = clustering_files)

# Display as a table
kable(clustering_df, caption = "Clustering Output Files (.rds)")
```
<br><br>

**Best WNN clustering result**: **Best WNN clustering result**: <mark>seurat.norm_counts_CRr_WNN_rnaHarm_atacHarm_k30_C.leiden_lsi_r2.rds</mark>

```{r}
library(here)

# Define the directory path
cluster_dir <- here("processed-data/05_Clustering_ARCr/01_clustering_std_method/seurat.norm_counts_CRr_WNN_rnaHarm_atacHarm_k30_C.leiden_lsi_r2.rds")
print(cluster_dir)
```

After evaluating the clustering results, we found that this configuration provides the most representative clusters for habenula. It captures the medial and lateral sub cell-types than other clustering settings, making it the most suitable choice for downstream analyses.
This is an object of class Seurat containing:
299552 features across <mark>55702</mark> samples (cells) within 2 assays 
Active assay: RNA (36601 features, 2000 variable features)
<br>
We have <mark>42 clusters</mark>, having <mark>13 of these with Habenula markers</mark>
<br><br>


**Habenula gene markers references**: (not styled)
<!-- ## code/04_DiffExpr_Clustering_seurat/remote_DGE_marker_gene_lists.R -->

<mark>Literature-Based + Data-Driven</mark>:

 [1] "DD_Astrocyte"                 "DD_Endo"                     
 [3] "DD_Excit.Thal"                "DD_Inhib.Thal"               
 [5] "DD_LHb"                       "DD_MHb"                      
 [7] "DD_Microglia"                 "DD_Oligo"                    
 [9] "DD_OPC"                       "LB_neuron"                   
[11] "LB_excitatory_neuron"         "LB_inhibitory_neuron"        
[13] "LB_Hb neuron specific"        "LB_MHB neuron specific"      
[15] "LB_LHB neuron specific"       "LB_oligodendrocyte"          
[17] "LB_oligodendrocyte_precursor" "LB_microglia"                
[19] "LB_astrocyte"                 "LB_Endo/CP"                  
[21] "LB_Thalamus broad"            "LB_Thalamus/MDm/Endo"        
[23] "LB_Thalamus/MDm/Endo-"        "LB_Thalamus/MDm"  

Total genes used: 469

<br><br>



**Habenula cell-types directory**: 
All DGE for all clusters directory:
processed-data/05_Clustering_ARCr/01_clustering_std_method/cvs_files/seurat.norm_counts_CRr_WNN_rnaHarm_atacHarm_k30_C.leiden_lsi_r2_markers.csv
DGE test: <mark>Wilcox test</mark>
Filtering steps: <mark>top 50 genes</mark> with <mark>FDR<0.05)</mark>
Top 50 DGE csv directory: 
processed-data/05_Clustering_ARCr/02_Hb_celltypes_from_seurat_reanalyze_v3/cvs_files_markers/WNN_rnaHarm_atacHarm_k30_C.leiden_lsi_r2_cellTypes_integrated_top50.csv"


```{r}
# Load knitr for kable
library(knitr)
library(here)

# Define the directory path
cluster_cell_types_dir <- here("processed-data/05_Clustering_ARCr/02_Hb_celltypes_from_seurat_reanalyze_v3")
print(cluster_dir)
# List only csv files
cell_types_files <- list.files(path = cluster_cell_types_dir, pattern = "\\.csv$", full.names = TRUE)

# Create a data frame for display
clustering_df <- data.frame(`RDS File` = cell_types_files)

# Display as a table
kable(clustering_df, caption = "Cell Types Files (.csv)")
```
<br><br>



**Seurat with cell-types annotated**: include all the clusters in the rna modality with labels with percentages of cells by cluster
```{r}
library(here)

# Define the directory path
cluster_cell_types_dir <- here("~/Hb_multiome/processed-data/05_Clustering_ARCr/05_rename_idents")
# List only .rds files
cell_types_files <- list.files(path = cluster_cell_types_dir, pattern = "\\.rds$", full.names = TRUE)
print(cell_types_files)

```

<br><br>


